---
title: "Play Animations"
author: "Amelie Devine"
date: "2024-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(gganimate)
library(sportyR)
library(nflplotR)
```


## read in data
```{r}
plays <- read_csv("plays.csv")

track_w2 <- read_csv("tracking_week_2.csv")
track_w3 <- read_csv("tracking_week_3.csv")
track_w6 <- read_csv("tracking_week_6.csv")
```

## select random plays
```{r}
filtered_plays <- plays %>% 
    filter(down == 1 | down == 2) %>% 
    filter(yardsToGo < 15 & yardsToGo > 4)

random <- filtered_plays %>% 
    sample_n(5) # selects new random each time
```

## filter for random plays
```{r}
sample_plays <- plays %>%  # these are 5 random initially selected that i like
    filter(gameId %in% c(2022092513, 2022091805, 2022091806, 2022101609, 2022101605) 
           & playId %in% c(736, 773, 3030, 1803, 1143)
           & possessionTeam %in% c('DEN', 'CAR', 'NE', 'SEA', 'NO'))

p1 <- track_season %>% 
    filter(gameId == 2022091805 & playId == 773) %>% 
    left_join(sample_plays, by = c("playId", "gameId"))

p2 <- track_season %>% 
    filter(gameId == 2022091806 & playId == 3030) %>% 
    left_join(sample_plays, by = c("playId", "gameId"))


p3 <- track_season %>% 
    filter(gameId == 2022092513 & playId == 736) %>% 
    left_join(sample_plays, by = c("playId", "gameId"))


p4 <- track_season %>% 
    filter(gameId == 2022101605 & playId == 1143) %>% 
    left_join(sample_plays, by = c("playId", "gameId"))


p5 <- track_season %>% 
    filter(gameId == 2022101609 & playId == 1803) %>% 
    left_join(sample_plays, by = c("playId", "gameId"))
```


```{r}
before_p1 <- p1 %>% 
    filter(frameType == "BEFORE_SNAP")

before_p2 <- p2 %>% 
    filter(frameType == "BEFORE_SNAP")

before_p3 <- p3 %>% 
    filter(frameType == "BEFORE_SNAP")

before_p4 <- p4 %>% 
    filter(frameType == "BEFORE_SNAP")

before_p5 <- p5 %>% 
    filter(frameType == "BEFORE_SNAP")

snap_p1 <- p1 %>% 
    filter(event == "ball_snap")

snap_p2 <- p2 %>% 
    filter(event == "ball_snap")

snap_p3 <- p3 %>% 
    filter(event == "ball_snap")

snap_p4 <- p4 %>% 
    filter(event == "ball_snap")

snap_p5 <- p5 %>% 
    filter(event == "ball_snap")
```


# try plotting play 1
```{r}
los_p1 <- snap_p1 %>% 
    filter(club=="football") %>% 
    select(x, y)

ggplot(snap_p1, aes(x=x, y=y, color=club)) +
    geom_point() +
    geom_vline(xintercept=los_p1$x, linetype="dotted") +
    labs(title = "Play 1 @ Snap of Ball",
         x = "X Position",
         y = "Y Position", 
         color = "Team") +
    scale_x_continuous(limits = c(0, 120),
                       breaks = seq(0, 120, by=10)) +
    scale_y_continuous(limits = c(0, 53.3)) +
    theme_minimal()
```

# try plotting play 2
```{r}
los_p2 <- snap_p2 %>% 
    filter(club=="football") %>% 
    select(x, y)

ggplot(snap_p2, aes(x=x, y=y, color=club)) +
    geom_point() +
    geom_vline(xintercept=los_p2$x, linetype="dotted") +
    labs(title = "Play 2 @ Snap of Ball",
         x = "X Position",
         y = "Y Position", 
         color = "Team") +
    scale_x_continuous(limits = c(0, 120),
                       breaks = seq(0, 120, by=10)) +
    scale_y_continuous(limits = c(0, 53.3)) +
    theme_minimal()
```

# try plotting play 3
```{r}
los_p3 <- snap_p3 %>% 
    filter(club=="football") %>% 
    select(x, y)

ggplot(snap_p3, aes(x=x, y=y, color=club)) +
    geom_point() +
    geom_vline(xintercept=los_p3$x, linetype="dotted") +
    labs(title = "Play 3 @ Snap of Ball",
         x = "X Position",
         y = "Y Position", 
         color = "Team") +
    scale_x_continuous(limits = c(0, 120),
                       breaks = seq(0, 120, by=10)) +
    scale_y_continuous(limits = c(0, 53.3)) +
    theme_minimal()
```


# try plotting play 4
```{r}
los_p4 <- snap_p4 %>% 
    filter(club=="football") %>% 
    select(x, y)

ggplot(snap_p4, aes(x=x, y=y, color=club)) +
    geom_point() +
    geom_vline(xintercept=los_p4$x, linetype="dotted") +
    labs(title = "Play 4 @ Snap of Ball",
         x = "X Position",
         y = "Y Position", 
         color = "Team") +
    scale_x_continuous(limits = c(0, 120),
                       breaks = seq(0, 120, by=10)) +
    scale_y_continuous(limits = c(0, 53.3)) +
    theme_minimal()
```


# try plotting play 5
```{r}
los_p5 <- snap_p5 %>% 
    filter(club=="football") %>% 
    select(x, y)

ggplot(snap_p5, aes(x=x, y=y, color=club)) +
    geom_point() +
    geom_vline(xintercept=los_p5$x, linetype="dotted") +
    labs(title = "Play 5 @ Snap of Ball",
         x = "X Position",
         y = "Y Position", 
         color = "Team") +
    scale_x_continuous(limits = c(0, 120),
                       breaks = seq(0, 120, by=10)) +
    scale_y_continuous(limits = c(0, 53.3)) +
    theme_minimal()
```

# What is the defense in???
```{r}
full_sample <- rbind(p1, p2, p3, p4, p5)

man <- full_sample %>% 
    filter(pff_manZone == "Man")

zone <- full_sample %>% 
    filter(pff_manZone == "Zone")

ggplot(full_sample, aes(x = pff_man))

```

## add markers for `event` and plot & save all plays
```{r}
plot1 <- ggplot(p1, aes(x = x, y = y, color = club)) +
            geom_rect(xmin = 0, xmax = 120, ymin = 0, ymax = 53.3,
                      fill = "darkgreen", color = "white", alpha = 0.2) +
            geom_vline(xintercept = seq(10, 110, by = 10), color = "white", linetype = "solid") +
            geom_segment(data = data.frame(x = rep(1:119, each = 2),
                                           y = rep(c(11, 42.3), times = 119)),
                         aes(x = x, xend = x, y = y, yend = y + 1), color = "white") +
            geom_point(aes(size = ifelse(is.na(nflId), 4, 2)), alpha = 0.8) +
            geom_text(aes(label = ifelse(is.na(nflId), "Football", jerseyNumber)),
                      vjust = -1, hjust = 0.5, color = "black", size = 3) +
            geom_text(aes(x = 60, y = 50, label = frameType), color = "gray", size = 6) +
            geom_text(aes(x = 10, y = 50, label = pff_manZone), color = "gray", size = 5) +
            scale_size_identity() +
            theme_minimal() +
            theme(legend.position = "none") +
            transition_states(frameId, transition_length = 2, state_length = 2, wrap = FALSE) +
            ease_aes('linear')

p1_anim <- animate(plot1, fps = 10, width = 800, height = 450, renderer = gifski_renderer())
anim_save("nfl_play1_animation_with_field.gif", p1_anim)

plot2 <- ggplot(p2, aes(x = x, y = y, color = club)) +
            geom_rect(xmin = 0, xmax = 120, ymin = 0, ymax = 53.3,
                      fill = "darkgreen", color = "white", alpha = 0.2) +
            geom_vline(xintercept = seq(10, 110, by = 10), color = "white", linetype = "solid") +
            geom_segment(data = data.frame(x = rep(1:119, each = 2),
                                           y = rep(c(11, 42.3), times = 119)),
                         aes(x = x, xend = x, y = y, yend = y + 1), color = "white") +
            geom_point(aes(size = ifelse(is.na(nflId), 4, 2)), alpha = 0.8) +
            geom_text(aes(label = ifelse(is.na(nflId), "Football", jerseyNumber)),
                      vjust = -1, hjust = 0.5, color = "black", size = 3) +
            geom_text(aes(x = 60, y = 50, label = frameType), color = "gray", size = 6) +
            geom_text(aes(x = 10, y = 50, label = pff_manZone), color = "gray", size = 5) +
            scale_size_identity() +
            theme_minimal() +
            theme(legend.position = "none") +
            transition_states(frameId, transition_length = 2, state_length = 2, wrap = FALSE) +
            ease_aes('linear')

p2_anim <- animate(plot2, fps = 10, width = 800, height = 450, renderer = gifski_renderer())
anim_save("nfl_play2_animation_with_field.gif", p2_anim)

plot3 <- ggplot(p3, aes(x = x, y = y, color = club)) +
            geom_rect(xmin = 0, xmax = 120, ymin = 0, ymax = 53.3,
                      fill = "darkgreen", color = "white", alpha = 0.2) +
            geom_vline(xintercept = seq(10, 110, by = 10), color = "white", linetype = "solid") +
            geom_segment(data = data.frame(x = rep(1:119, each = 2),
                                           y = rep(c(11, 42.3), times = 119)),
                         aes(x = x, xend = x, y = y, yend = y + 1), color = "white") +
            geom_point(aes(size = ifelse(is.na(nflId), 4, 2)), alpha = 0.8) +
            geom_text(aes(label = ifelse(is.na(nflId), "Football", jerseyNumber)),
                      vjust = -1, hjust = 0.5, color = "black", size = 3) +
            geom_text(aes(x = 60, y = 50, label = frameType), color = "gray", size = 6) +
            geom_text(aes(x = 10, y = 50, label = pff_manZone), color = "gray", size = 5) +
            scale_size_identity() +
            theme_minimal() +
            theme(legend.position = "none") +
            transition_states(frameId, transition_length = 2, state_length = 2, wrap = FALSE) +
            ease_aes('linear')

p3_anim <- animate(plot3, fps = 10, width = 800, height = 450, renderer = gifski_renderer())
anim_save("nfl_play3_animation_with_field.gif", p3_anim)

plot4 <- ggplot(p4, aes(x = x, y = y, color = club)) +
            geom_rect(xmin = 0, xmax = 120, ymin = 0, ymax = 53.3,
                      fill = "darkgreen", color = "white", alpha = 0.2) +
            geom_vline(xintercept = seq(10, 110, by = 10), color = "white", linetype = "solid") +
            geom_segment(data = data.frame(x = rep(1:119, each = 2),
                                           y = rep(c(11, 42.3), times = 119)),
                         aes(x = x, xend = x, y = y, yend = y + 1), color = "white") +
            geom_point(aes(size = ifelse(is.na(nflId), 4, 2)), alpha = 0.8) +
            geom_text(aes(label = ifelse(is.na(nflId), "Football", jerseyNumber)),
                      vjust = -1, hjust = 0.5, color = "black", size = 3) +
            geom_text(aes(x = 60, y = 50, label = frameType), color = "gray", size = 6) +
            geom_text(aes(x = 10, y = 50, label = pff_manZone), color = "gray", size = 5) +
            scale_size_identity() +
            theme_minimal() +
            theme(legend.position = "none") +
            transition_states(frameId, transition_length = 2, state_length = 2, wrap = FALSE) +
            ease_aes('linear')

plot4

p4_anim <- animate(plot4, fps = 10, width = 800, height = 450, renderer = gifski_renderer())
anim_save("nfl_play4_animation_with_field.gif", p4_anim)

plot5 <- ggplot(p5, aes(x = x, y = y, color = club)) +
            geom_rect(xmin = 0, xmax = 120, ymin = 0, ymax = 53.3,
                      fill = "darkgreen", color = "white", alpha = 0.2) +
            geom_vline(xintercept = seq(10, 110, by = 10), color = "white", linetype = "solid") +
            geom_segment(data = data.frame(x = rep(1:119, each = 2),
                                           y = rep(c(11, 42.3), times = 119)),
                         aes(x = x, xend = x, y = y, yend = y + 1), color = "white") +
            geom_point(aes(size = ifelse(is.na(nflId), 4, 2)), alpha = 0.8) +
            geom_text(aes(label = ifelse(is.na(nflId), "Football", jerseyNumber)),
                      vjust = -1, hjust = 0.5, color = "black", size = 3) +
            geom_text(aes(x = 60, y = 50, label = frameType), color = "gray", size = 6) +
            geom_text(aes(x = 10, y = 50, label = pff_manZone), color = "gray", size = 5) +
            scale_size_identity() +
            theme_minimal() +
            theme(legend.position = "none") +
            transition_states(frameId, transition_length = 2, state_length = 2, wrap = FALSE) +
            ease_aes('linear')

p5_anim <- animate(plot5, fps = 10, width = 800, height = 450, renderer = gifski_renderer())
anim_save("nfl_play5_animation_with_field.gif", p5_anim)
```



# plot best & worst plays
```{r, include=FALSE}
track_season <- track_season %>% 
    mutate(gamePlayId = paste0(gameId, playId))
```

#### lowest prob, true negative: 20221009033841 (pred zone, actual zone)
```{r}
tn_low <- track_season %>% 
    filter(gamePlayId == 20221009033841) %>% 
    left_join(plays, by = c("playId", "gameId"))

ggplot(tn_low, aes(x = x, y = y, color = club)) +
            geom_rect(xmin = 0, xmax = 120, ymin = 0, ymax = 53.3,
                      fill = "darkgreen", color = "white", alpha = 0.2) +
            geom_vline(xintercept = seq(10, 110, by = 10), color = "white", linetype = "solid") +
            geom_segment(data = data.frame(x = rep(1:119, each = 2),
                                           y = rep(c(11, 42.3), times = 119)),
                         aes(x = x, xend = x, y = y, yend = y + 1), color = "white") +
            geom_point(aes(size = ifelse(is.na(nflId), 4, 2)), alpha = 0.8) +
            geom_text(aes(label = ifelse(is.na(nflId), "Football", jerseyNumber)),
                      vjust = -1, hjust = 0.5, color = "black", size = 3) +
            geom_text(aes(x = 60, y = 50, label = frameType), color = "gray", size = 6) +
            geom_text(aes(x = 10, y = 50, label = pff_manZone), color = "gray", size = 5) +
            geom_text(aes(x = 110, y = 50, label = event), color = "gray", size = 5) +
            scale_size_identity() +
            theme_minimal() +
            theme(legend.position = "none") +
            transition_states(frameId, transition_length = 4, state_length = 4, wrap = FALSE) +
            ease_aes('linear')
```

#### highest prob, true positive: 20221010001500 (pred man, actual man)
```{r}
tp_high <- track_season %>% 
    filter(gamePlayId == 20221010001500) %>% 
    left_join(plays, by = c("playId", "gameId"))

ggplot(tp_high, aes(x = x, y = y, color = club)) +
            geom_rect(xmin = 0, xmax = 120, ymin = 0, ymax = 53.3,
                      fill = "darkgreen", color = "white", alpha = 0.2) +
            geom_vline(xintercept = seq(10, 110, by = 10), color = "white", linetype = "solid") +
            geom_segment(data = data.frame(x = rep(1:119, each = 2),
                                           y = rep(c(11, 42.3), times = 119)),
                         aes(x = x, xend = x, y = y, yend = y + 1), color = "white") +
            geom_point(aes(size = ifelse(is.na(nflId), 4, 2)), alpha = 0.8) +
            geom_text(aes(label = ifelse(is.na(nflId), "Football", jerseyNumber)),
                      vjust = -1, hjust = 0.5, color = "black", size = 3) +
            geom_text(aes(x = 60, y = 50, label = frameType), color = "gray", size = 6) +
            geom_text(aes(x = 10, y = 50, label = pff_manZone), color = "gray", size = 5) +
            geom_text(aes(x = 110, y = 50, label = event), color = "gray", size = 5) +
            scale_size_identity() +
            theme_minimal() +
            theme(legend.position = "none") +
            transition_states(frameId, transition_length = 4, state_length = 4, wrap = FALSE) +
            ease_aes('linear')

## LB clearly follows WR in motion
```

#### highest prob, false positive: 2022101606340 (pred man, actual zone)
```{r}
fp_high <- track_season %>% 
    filter(gamePlayId == 2022101606340) %>% 
    left_join(plays, by = c("playId", "gameId"))

ggplot(fp_high, aes(x = x, y = y, color = club)) +
            geom_rect(xmin = 0, xmax = 120, ymin = 0, ymax = 53.3,
                      fill = "darkgreen", color = "white", alpha = 0.2) +
            geom_vline(xintercept = seq(10, 110, by = 10), color = "white", linetype = "solid") +
            geom_segment(data = data.frame(x = rep(1:119, each = 2),
                                           y = rep(c(11, 42.3), times = 119)),
                         aes(x = x, xend = x, y = y, yend = y + 1), color = "white") +
            geom_point(aes(size = ifelse(is.na(nflId), 4, 2)), alpha = 0.8) +
            geom_text(aes(label = ifelse(is.na(nflId), "Football", jerseyNumber)),
                      vjust = -1, hjust = 0.5, color = "black", size = 3) +
            geom_text(aes(x = 60, y = 50, label = frameType), color = "gray", size = 6) +
            geom_text(aes(x = 10, y = 50, label = pff_manZone), color = "gray", size = 5) +
            geom_text(aes(x = 110, y = 50, label = event), color = "gray", size = 5) +
            scale_size_identity() +
            theme_minimal() +
            theme(legend.position = "none") +
            transition_states(frameId, transition_length = 4, state_length = 4, wrap = FALSE) +
            ease_aes('linear')

## LB follows WR in motion
```

#### lowest prob, false negative: 20221023032512 (pred zone, actual man)
```{r}
fn_low <- track_season %>% 
    filter(gamePlayId == 20221023032512) %>% 
    left_join(plays, by = c("playId", "gameId"))

ggplot(fn_low, aes(x = x, y = y, color = club)) +
            geom_rect(xmin = 0, xmax = 120, ymin = 0, ymax = 53.3,
                      fill = "darkgreen", color = "white", alpha = 0.2) +
            geom_vline(xintercept = seq(10, 110, by = 10), color = "white", linetype = "solid") +
            geom_segment(data = data.frame(x = rep(1:119, each = 2),
                                           y = rep(c(11, 42.3), times = 119)),
                         aes(x = x, xend = x, y = y, yend = y + 1), color = "white") +
            geom_point(aes(size = ifelse(is.na(nflId), 4, 2)), alpha = 0.8) +
            geom_text(aes(label = ifelse(is.na(nflId), "Football", jerseyNumber)),
                      vjust = -1, hjust = 0.5, color = "black", size = 3) +
            geom_text(aes(x = 60, y = 50, label = frameType), color = "gray", size = 6) +
            geom_text(aes(x = 10, y = 50, label = pff_manZone), color = "gray", size = 5) +
            geom_text(aes(x = 110, y = 50, label = event), color = "gray", size = 5) +
            scale_size_identity() +
            theme_minimal() +
            theme(legend.position = "none") +
            transition_states(frameId, transition_length = 4, state_length = 4, wrap = FALSE) +
            ease_aes('linear')

## CB follows a little but not much
```

